2.2-1 \\
Express the function $n^3/1000 + 100n^2 - 100n + 3$ in terms of $\theta$-notation. \\

So we have $c_1n^3 + c_2n^2 - c_3n$ we know the we can cancel out the constant terms since
they are negligable for large vlaues of n and we know $n^3$ grows the fastest since 
it is a cubic. $\theta(n^3)$ is the answer \\

2.2-2 \\
Consider sorting n numbers stored in array A[1:n] by first finding the smallest element of A[1:n]
and exchanging it with the element in A[1]. Then find the smalles element of A[2:n], and exchange it with A[2]. 
Then find the smallest element of A[3:n], and exchange it with A[3]. Continue this for the first n-1 elements of A.
Write the code for this algorithm (this is selection sort). What loop invariant does this algorithm maintain? Why does it need
to run for only the first n-1 elements, rantehr than all n? Give the worst-case running time of selection sort in $\theta -notation$. Is the best-case
run time any better? \\

\begin{verbatim}
# selection sort code (in place)
def selectionSort(array: list[int]) -> None:
    MAX_INT: int = 1000000000
    
    n: int = len(array)
    i: int = 0
    
    while (i < n-1):
        cur_min: int = MAX_INT
        min_i: int = 0
        
        for j in range(i, n):
            num: int = array[j]
            
            if (num < cur_min):
                cur_min = num
                min_i = j
        
        temp: int = array[i]
        array[i] = array[min_i]
        array[min_i] = temp
        i += 1
\end{verbatim}

The loop invariant that this algorithm maintains (the property that holds true before and after every iteratoin of the loop).
A[0:i-1] contains the i-1 smallest elements of the array and they are in sorted order \\

The algorithm only needs to run for the first n-1 elements because of the way it works
it selects the smallest element for the current subarray being checked A[i:n] it then places that in the lowest postion 
in subarray and does this $\forall i \in [0:n-1]$ so by the time i gets to n-1 this sub array is A[n-1:n] which only contains
1 element and thus is already sorted. \\

The worst case runtime of selection sort would be $\theta(n^2)$ \\

The best case runtime of selection sort would be $\theta(n^2)$ which is the same. There is no mechanism to check if the subarray
being checked is already sorted.\\

2.2-3 \\
Consider linear search again (see Exercise 2.1-4). How many elements of the input 
array need to be checked on the average, assuming that the element being searched 
for is equally likely to be any element in the array? How about in the worst case? 
Using $\theta -notation$, give the average-case and worst-case run times of linear 
search. Justify your answers. \\

Let x be the element we are searching for. 
Let n be the number of elements in the array. \\

Assuming that x exists in the array then statistically on average we should have to 
check roughly n/2 elements to find x.  
Since x is equally likely to be at each index the average would look like $\frac{(1 + 2 + ... + n)}{n}$
and we know $1 + 2 + ... + n = \frac{n(n+1)}{2}$ thus using substitution we get an average runtime of
$\frac{n(n+1)}{2n} = \frac{n+1}{2} = \frac{n}{2} + \frac{1}{2}$\\

In the worst cast the element would be at 
the last index in the array so at A[n-1] in this scenerio we have to check all n elements to find x. \\

Notice: the dominant term for both the average and worst case remains $\theta(n)$ the only thing that changes
is the constants. \\

2.2-4 \\
How can you modify any sorting algorithm to have a good best-case running time? \\
You can add a way to check if the subarray you are currently finding the min for is already sorted. \\ 
so maybe somthing like this \\

\begin{verbatim}
# selection sort code (in place)
def selectionSort(array: list[int]) -> None:
    
    n: int = len(array)
    i: int = 0
    
    while (i < n-1):
        cur_min: int = array[i]
        min_i: int = 0
        already_sorted: bool = True
        
        for j in range(i, n):
            num: int = array[j]
            
            if (j != n-1 and array[j + 1] < num):
                already_sorted = False
                
            if (num < cur_min):
                cur_min = num
                min_i = j
        
        if (already_sorted):
            return
        
        temp: int = array[i]
        array[i] = array[min_i]
        array[min_i] = temp
        i += 1
\end{verbatim}


