2.1-4 \\
Consider the searching problem: \\
\textbf{Input}: A sequence of \textit{n} numbers $<a_1, a_2, ..., a_n>$ stored in array A[1:n] and a value x. \\
\textbf{Output}: An index \textit{i} that \textit{x} equals A[i] or the special value \textit{NIL} if \textit{x} doesn't appear in A. \\

Write pseudocode for linear search, which scans through the array from begin-ning to end, looking for 
x. Using a loop invariant, prove that your algorithm is 
correct. Make sure that your loop invariant fulûlls the three necessary properties.

\begin{verbatim}
# give the index of the x target or else -1 if not found
def find_val(x: int, array: list[int]) -> int:
    n: int = len(array)
    res: int = -1 

    for i in range(n):
        if (array[i] == x):
            res = i
            break
            
    return res
            
array: list[int] = [1,2,3,4,5,6,7]
index: int = find_val(100, array)

print(f"Index: {index}")
\end{verbatim}

Note: Loop invariant properites include:
\begin{itemize}
    \item \textbf{Initialization}: The invariant is true prior to the first iteration of the loop.
    
    \item \textbf{Maintenance}: If the invariant is true before an iteration of the loop, it remains true before the next iteration.
    
    \item \textbf{Termination}: The loop eventually terminates. When it does, the invariant—together with the reason the loop terminated—gives a useful property that helps prove the algorithm is correct.
\end{itemize}

\begin{proof} 

\textbf{Initialization:} Before the first iteration, the loop variable $i$ is set to $0$.
At this point, the subarray $A[0:i]$ contains no elements. Therefore, no index $j < i$
exists such that $A[j] = x$. The loop invariant holds because we have not yet examined
any elements, so it is correct that $x$ has not been found in the subarray. Thus we return
Nil (-1) which is a valid solution in this scenerio. \\

\textbf{Maintenance}: Next, we show that each iteration maintains the loop invariant. This means that if the invariant holds for the 
nth iteration, then it also holds for the (n+1)th iteration \\

Assume that at the start of the nth iteration the invariant holds. That is, we have already checked all indices 
$i \in [0,n-1]$ and found that $A[i] \neq x$. Therefore the value x has not been found in the first n element. \\ 

During the next iteration we check at index n. If A[n] == x then the algorithm returns the index n. Otherwise we continue
to the next iteration and we know that A[0:n] does not conatin x \\

Thus loop invariant holds for the next iteration. And thus we have shown Maintenance holds for the algorithm. 

\textbf{Termination} The loop terminates when either the element x is found or when the index i reaches n,
the lengthe of the array. Since the loop increments i each iteration and $0 \leq i \leq n$, the algorithm must terminate. \\

If the algorithm terminate because A[i] == x, then the it correctly returns the index i. \\

If the loop terminates because i == n, the by the loop invariant we know that x is not in A[0:n-1]. Therefore, the element x does not appear in the array,
and we return -1 indicating this. 

\end{proof}

2.1-5 \\
Consider the problem of adding two n-bit binary integers a and b, stored in two n-element arrays
A[0:n-1] and B[0:n-1], where each element is either 0 or 1, a = $\sum_{i= 0}^{n-1}$A[i]*$2^i$, and b = $\sum_{i=0}^{n-1}B[i]*2^i$.
The sum c = a + b of the two integers should be stored in binary form in an (n + 1)-element array C[0:n], where c = $\sum_{i=0}^{n}C[i]*2^i$.
Write a procedure ADD-BINARY-INTEGERS that takes as input arrays A and B, along with the length n, and returns array C holding the sum. \\

\begin{verbatim}
    def add_binary_arrays(binary1: list[str], binary2: list[str]) -> list[str]:
    # keep pointers for binary 1 and binary 2
    # from the question they have the same len so 
    # it is fair to say len(binary1) == len(binary2) == n
    n: int = len(binary1)
    i: int = n-1
    carry: bool = False
    C: list[str] = ['0'] * (n + 1)
    
    
    while (i >= 0):
        bit1: str = binary1[i]
        bit2: str = binary2[i]
        
        if (bit1 == '1' and bit2 == '1'):
            if (carry):
                C[i+1] = '1'
            carry = True
            
        elif ((bit1 == '1' and bit2 == '0') 
            or (bit1 == '0' and bit2 == '1')):
            if (not carry):
                C[i+1] = '1'
                carry = False
            else:
                carry = True
                
        else: # both are 0
            if (carry):
                C[i+1] = '1'
            carry = False
            
        i -= 1
    
    # check carry for the last bit
    if (carry):
        C[0] = '1'
    
    return C
\end{verbatim}

